---
title: "Jackpot Size"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{jackpot_size}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lottodata)
```

## lottodata

The get access to this data set, simply, install the `lottodata` package via [GitHub](https://github.com/andr3wli/lottodata) and type `jackpot_size` into R. 

## `jackpot_size`

The variables included in the data set:

| Variable       | Description                                                        | Type of variable |
|----------------|--------------------------------------------------------------------|------------------|
| `zip_code`     | The first 3 digits of postal code (geographical region)            | string           |
| `start_date`   | The start of the sales date (year-month-day format)                | date             |
| `end_date`     | The end of the sales date (year-month-day format)                  | date             |
| `game`         | The specific lottery game (one of: Lotto Max, Lotto 649, Lottario) | string           |
| `ticket_sales` | Number of tickets sold                                             | integer          |
| `net_sales`    | The total cad dollar amount of sales                               | integer          |
| `jackpot_size` | The jackpot size in cad dollars                                    | integer          |
| `year`         | Year                                                               | integer          |
| `month`        | Month                                                              | integer          |
| `day`          | Day                                                                | integer          |


## Example
Suppose you want to look at the game Lottario more in 2014:
```{r}
library(lottodata)
library(dplyr)

# What is the yearly spending for the Lottario in zone M1B in 2012?
jackpot_size %>%
  filter(year == 2012 & game == "Lottario" & zip_code == "M1B") %>%
  head()
```

## ggplot2 example
```{r}
library(ggplot2)
theme_set(theme_classic())

jackpot_data <- jackpot_size
jackpot_plot <- jackpot_data %>%
  filter(year == 2012 & game == "Lottario" & zip_code == "M1B") %>%
  ggplot(aes(day, ticket_sales, fill = as.factor(month))) +
  geom_col() +
  facet_wrap(~month, labeller = labeller(month = 
                                           c("1" = "January", "2" = "February", "3" = "March", "4" = "April", "5" = "May",
                                             "6" = "June", "7" = "July", "8" = "August", "9" = "September", "10" = "October",
                                             "11" = "November", "12" = "December"))) +
  labs(x = "Days", y = "# of tickets sold", title = "Lottario ticket salees in 2012") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("#a6cee3","#1f78b4","#b2df8a","#33a02c","#fb9a99","#e31a1c",
                               "#fdbf6f","#ff7f00","#cab2d6","#6a3d9a","#ffff99","#b15928"))

jackpot_plot
```


## Example exploratory data analysis:
```{r}
# EDA via base R

jackpot_eda <- function(x){
  hist(x, col = rainbow(30))
  plot(x)
  plot(density(x))
  data.frame(min = min(x),
             median = median(x),
             mean = mean(x),
             max = max(x),
             sd = sd(x),
             range =max(x) - min(x) )
}

jackpot_eda(jackpot_size$ticket_sales)
```

