---
title: "Demographics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lotto_demographics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lottodata)
```

## lottodata
The get access to this data set, simply, install the `lottodata` package via [GitHub](https://github.com/andr3wli/lottodata) and type `lotto_demograpics` into R. 


## `lotto demographics`
The variables included in the data set:

| Variable      | Description                                                                                                                                                                                                                                                                                                                                                                                                            | Type of variable |
|---------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|------------------|
| `zip_code`    | The first 3 digits of postal code (geographical region)                                                                                                                                                                                                                                                                                                                                                                | string           |
| `geo_id`      | Geography ID                                                                                                                                                                                                                                                                                                                                                                                                           | integer          |
| `income`      | Per capita income levels                                                                                                                                                                                                                                                                                                                                                                                               | integer          |
| `education`   | Highest completed level of education for the population                                                                                                                                                                                                                                                                                                                                                                | float            |
| `mbsa`        | Proportion of time spent in white collar employment. White collar <br>employment is defined as the proportion of residents aged 15 or <br>greater employed in management, business finance and administration, <br>health, education, law, social community and government services, <br>art, culture, natural and applied sciences and related occupations, <br>according to the National Occupational Classification | float            |
| `ses`         | SES was calculated via takling the sum of the Z-scores of it's <br>per-capita income, years of education, and proportion of white-collar <br>workers                                                                                                                                                                                                                                                                   | float            |
| `description` | Describes where the location is in natural language                                                                                                                                                                                                                                                                                                                                                                    | string           |


## Example exploratory data analysis:
```{r}
# EDA via base R

demographics_eda <- function(x){
  hist(x, col = rainbow(30))
  plot(x)
  plot(density(x))
  data.frame(min = min(x),
             median = median(x),
             mean = mean(x),
             max = max(x),
             sd = sd(x),
             range =max(x) - min(x) )
}

demographics_eda(lotto_demographics$income)
```
